---
title: "Class 5 practice - Planetscope"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Instructions
Execute all chunks separately. i.e. ctrl-shift-enter while the cursor is in the chunk. 

Make sure this returns your file path ending in "geog287387/". If it doesn't, uncomment out line 19 and edit it match your path to the geog287387 project.  
```{r, eval=FALSE}
bpath <- paste0(getwd(), "/")
# bpath <- "D:/Users/<your_user_name>/Documents/geog287387/"
bpath
```


## Packages
```{r, eval=FALSE}
# setwd("..")
library(gdalUtils)
library(raster)
script_path <- paste0(bpath, "materials/code/R/uas_image_functions.R")
source(script_path)
```

## Paths
Set up paths to downloaded Planetscope imagery, and create output directory for adjusting images. You can make notes here. 
```{r, eval=FALSE}
img_path <- paste0(bpath, "materials/data/imagery/planet/")
out_dir <- paste0(bpath, "materials/data/05")

if(!dir.exists(out_dir)) dir.create(out_dir)

planet_dirs <- dir(img_path, pattern = "2019", full.names = TRUE)
planet_imgs <- lapply(planet_dirs, function(x) {
  dir(x, pattern = "MS_SR", full.names = TRUE)
})
```

## Read in image details

Testing read in and plot of one image
```{r, eval=FALSE}
b1 <- brick(planet_imgs[[1]])

# Use gdal to reproject images to EPSG4326 and to a defined extent
ext <- c(-71.8101, 42.1193, -71.8041, 42.1242)
out1 <- substr(basename(planet_imgs[[1]]), 1, 8)
prj <- proj4string(b1)
planet1 <- img_align(img = b1@file@name, res = 3 / 100000, out_dir = out_dir, 
                     out_name = out1, ext = ext)
plotRGB(planet1[[c(4, 3, 2)]], stretch = "lin")
```

## Read in and process all PlanetScope images
```{r, eval=FALSE}
ext <- c(-71.8101, 42.1193, -71.8041, 42.1242)

plist <- lapply(1:length(planet_imgs), function(x) {
  nm <- planet_imgs[[x]]
  b <- brick(nm)
  out <- paste0(substr(basename(nm), 1, 8), "_", x)
  prj <- proj4string(b)
  planet1 <- img_align(img = b@file@name, res = 3 / 100000, out_dir = out_dir,                        out_name = out, ext = ext)
})
```

## Drone image
Read in the August 26, 2019 UAS orthomosaic processed during class 4. It should have been processed with PPK correction and against the reflectance target.  
```{r, eval=FALSE}
uas <- brick("materials/data/04/aug262019_ngb_cog_gcs.tif")
```

## Extract time series from imagery
### Create random points
And use them to first collect points from the drone data, and filter out those pts falling in NA areas of drone data
```{r, eval=FALSE}
library(sf)
library(dplyr)

# create random sample of spatial points
set.seed(1)
pts <- sf::st_sample(st_as_sf(as(extent(uas[[1]]), "SpatialPolygons")), 
                     size = 1000, exact = TRUE)
pts <- cbind(ID = 1:length(pts), st_sf(pts))  # convert points to sfc

# use points to extract values from uas raster
uas_dat <- extract(uas, y = pts)
uas_dat <- cbind(uas_dat, 
                 ndvi = (uas_dat[, 1] - uas_dat[, 2]) / 
                   (uas_dat[, 1] + uas_dat[, 2])) 
pts <- cbind(pts, uas_dat) 
colnames(pts) <- c("ID", "nir", "red", "green", "ndvi", "pts")
plotRGB(uas, stretch = "lin")
plot(st_geometry(pts), add = TRUE)

# then reduce pts to just those falling in UAS imaged area
pts <- pts %>% filter(!is.na(nir))
plot(st_geometry(pts), add = TRUE, col = "green", pch = 16) # check their distr

```

## Time series

Now use points to extract points from Planet time series, but first
```{r, eval = FALSE}
# create vector of dates
dates <- substr(basename(unlist(planet_imgs)), 1, 8)
dates <- lubridate::as_date(dates)

# check images to see which have complete coverage
par(mfrow = c(3, 4), mar = rep(3, 4))
for(i in 1:length(plist)) {
  plot(plist[[i]][[4]], main = dates[i])
}

# select just full coverage images
plist2 <- plist[c(3, 4, 7, 8, 9, 10, 12)]
dates2 <- dates[c(3, 4, 7, 8, 9, 10, 12)]

# Note: (a fancier way to do this would be to count NAs in each image), e.g. 
# which(sapply(plist, function(x) length(which(is.na(values(x[[4]]))))) == 0)
# but you would have to check by date as well (this makes it more complicated,
# here is how you would do it below -- not run)
# plist2 <- lapply(unique(dates), function(x) {  # iterate over unique dates
#   plist_1dy <- plist[dates == x]  # pull out images for just that date
#   ind <- sapply(plist_1dy, function(y) {
#     length(which(is.na(values(y[[4]]))))
#   })  # vector of length of NAs in each NIR band of each image from this date
#   # randomly select image w/no nas
#   set.seed(1)
#   out <- plist_1dy[[sample(which(ind == 0), 1)]]
#   return(out)
# })

# extract points from all rasters
ppts <- lapply(plist2, function(x) raster::extract(x, y = pts))

# subset to NIR, red, green, blue, ndvi
nir <- lapply(ppts, function(x) x[, 4])
red <- lapply(ppts, function(x) x[, 3])
green <- lapply(ppts, function(x) x[, 2])
blue <- lapply(ppts, function(x) x[, 1])
ndvi <- lapply(ppts, function(x) (x[, 4] - x[, 3]) / (x[, 4] + x[, 3]))

# calculate mean of NIR NDVI
nir_mu <- sapply(nir, function(x) mean(x, na.rm = TRUE))
ndvi_mu <- sapply(ndvi, function(x) mean(x, na.rm = TRUE))

## mean of times series
dev.off() # resets previous plots
plot(dates2, nir_mu, main = "NIR mean", type = "l")

# ndvi mean
plot(dates2, ndvi_mu, type = "l")

# select 10 random points from images and look at them
set.seed(40)
ss <- sample(1:nrow(pts), size = 10, replace = FALSE)

# use it to select from the already extracted points
ndvi_ss <- lapply(ndvi, function(x) x[ss])  # ndvi
ndvi_ss <- do.call(rbind, ndvi_ss)  # bind into matrix
nir_ss <- lapply(nir, function(x) x[ss])  # nir
nir_ss <- do.call(rbind, nir_ss) # bind into matrix

# plot
plot(dates2, ndvi_ss[, 1], ylim = c(0, 1), type = "l")  # first date first
for(i in 2:10) lines(dates2, ndvi_ss[, i])  # 

plot(dates2, nir_ss[, 1], ylim = c(0, 1), type = "l")
for(i in 2:10) lines(dates2, nir_ss[, i] / 10000)  # nir/10000 to scale to refl

# where are these points?
plot(plist2[[1]][[1]])
plot(st_geometry(pts[ss, ]), add = TRUE)
```

## UAS-Planet comparison
Now compare, for those same 10 points, the data from the drone to the values from the Planet image collected on the closest date (August 24) to see how different they are. Compare the NIR and red channels as well NDVI. The code above provides everything you need to do that. Note that we have already extracted all the values from the UAS for the full set of sample points. We also have already extracted from the PlanetScope data, and subset the values. You just need to select the right subset of points to compare from the UAS data to compare. The comparison is now on a scatter plot. I'll get you started.

```{r, eval=FALSE}
uas_nir <- pts %>% slice(ss) %>% pull(nir)  # pulls the values out pts for nir 
plot(uas_nir, nir_ss[6, ] / 10000, xlab = "UAS NIR", ylab = "Planet NIR") # plot
```

We select the 6th row of `nir_ss` because the rows in the matrix represent different image dates (columns are different points from the image), and August 24th is the 6th or 7 PlanetScope images in the time series. 


```{r, eval=FALSE}
uas_ndvi <- pts %>% slice(ss) %>% pull(ndvi) # pulls the values out pts for nir 
plot(uas_ndvi, ndvi_ss[6, ], xlab = "UAS NIR", ylab = "Planet NIR") # plot
```

Now do the same for red and ndvi. 

## Scaling comparison
Now, using the code from the previous class:

- aggregate the UAS data to the same resolution as the PlanetScope data. Use a factor of 27 to aggregate
- Create a file `pts2` using this argument `pts2 <- pts %>% select(ID)`
- then, using all of `pts2`, extract the NDVI, NIR, red, and green reflectance values from aggregated drone data. Lines 95-100 will help you recreate the steps you need to extract from aggregated uas data using `pts`. 
- Once extract, compare aggregated UAS reflectances (NDVI, NIR, red, and green) to those from the August 24 Planet (which doesn't need aggregation).  
by plotting the extracted points for each band against the relevant PlanetScope subsets from lines 147-151 above. Use the correct PlanetScope data (image 6 in the time series). Note that red, green, and NIR needed to be scaled by 10000. 
- The code from class 4 practice can show you how to plot the values in a scatter plot with the 1:1 line. 
- Be ready to show your results in class 
```{r, eval=FALSE}
# code here
```









