---
title: "Cloud-based Processed => Realtime In Situ Sensing"
subtitle: "New Methods in Earth Observation"
# author: "Lyndon Estes"
date: "10/3/2018"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "lucy", "middlebury-fonts", "themes/class11-theme.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: center, middle

# Cloud Processing Practicals/Homework
## Wrap-up

---
## Earth Engine 
### 1. GCVI on Landsat

[Code](https://code.earthengine.google.com/21f7df3222b920bc725ab242c729ec37)

### 2. Upload Planet and UAS imagery

### 3. Planet ImageCollection and GCVI

[Code](https://code.earthengine.google.com/21f7df3222b920bc725ab242c729ec37)

---
## Radiant Earth
[Login](https://app.radiant.earth)


---
## Working with Arable Mark data

Get the `aRable` R package .tgz from Slack (Mac), or the .zip file for Windows. 
First we will go through the install steps. Run the code below in a new source (script) file in Rstudio

### Windows
```{r, eval=FALSE}
setwd("c:/Users/airg/Desktop/")  # change the path to where you have zip
install.packages("devtools")  # if you don't already have it -->
install.packages("tidyverse")
install.packages("data.table")
install.packages("gridExtra")
install.packages("aRable.zip",  repos = NULL, type = "binary")
```

### Mac
Same as above, but change the first and last lines as below
```{r, eval=FALSE}
setwd("~/Desktop/")  # change the path to where you have zip
install.packages("aRable_0.1.1.tgz", repos = NULL)
```
---

## Load the Arable Package with Helpers

```{r, message=FALSE, warning=FALSE}
library(aRable)
library(data.table)
library(ggplot2)
library(gridExtra)
```

That loads the Arable client. Let's look at it's help file
```{r, eval=FALSE}
?ArableClient
```

---

## Let's use the Client

There are a couple of things to note: 

##### 1. We have three devices--their IDs are A000667, A000671,  and A000680
##### 2. For some reason, the client fails if you enter some of the parameter values in separate variables. So please enter values directly in the client. 

We are going to start by selecting daily variables for device A000667 for the month of June

**Note that I have the value "<password>" specified below. Replace that with the password I sent separately via Slack. **
```{r, eval = FALSE}
dev1 <- ArableClient(device = "A000667", measure = "daily", 
                     start = "2018-06-01", end = "2018-06-30", 
                     email = "lestes@clarku.edu", 
                     password = "<password>",  # replace this w/ actual p
                     tenant = "clark")
# a little cleaning
dev1 <- data.table(dev1)
dev1[is.na(precip), precip := 0]  # set NA precip to 0
dev1 <- na.omit(dev1)  # remove missing dates
```

```{r, echo=FALSE, eval=FALSE}
save(dev1, file = "materials/data/12/dev1.rda")
```
```{r, echo=FALSE}
# getwd()
load("../data/12/dev1.rda")
```

---

That collected some data, and now we can have a look at it.
```{r, warning=FALSE, fig.width=7, fig.height=6, fig.align='center'}
dat <- dev1  # change this when you change devices
mytheme <- function() theme(axis.title.x = element_blank())
p1 <- ggplot(dat) + geom_line(aes(datetime, precip), col = "blue") + mytheme()
p2 <- ggplot(dat) + geom_line(aes(datetime, NDVI), col = "green3") + mytheme()
p3 <- ggplot(dat) + geom_line(aes(datetime, meanT), col = "purple") + mytheme()
p4 <- ggplot(dat) + geom_line(aes(datetime, CGDD), col = "orange") + mytheme()
grid.arrange(p1, p2, p3, p4, ncol = 2)
```

---

## Assignment

**Part 1: Reuse the code above to do the following:**

- Change the date range to June 1 to September 6, and rerun the downloads from the Arable API
  - i.e. Change `end` argument to "2018-09-06"
- Do that for each of the three devices
  - Change the `device` argument between "A000667", "A000671", "A000680"
  - direct the output to `dev1`, `dev2`, `dev3`
- Reuse the plot code for `dev1`, `dev2`, `dev3`
  - i.e. copy and paste the code block from slide 8
  - replace `dev1` with `dev2` or `dev3` in the `dat` argument, e.g. `dat <- dev2`
- This means you will have three blocks of download code, for `dev1`, `dev2`, `dev3`, and three blocks of plotting code for `dev1`, `dev2`, `dev3`

---

- Extra! You may wish to explore other variables 

  - You can do that by substituting different variable names contained in `dev1`, `dev2`, or `dev3` into the plotting code. 

  - E.g. replace `CGDD` in this `geom_line(aes(datetime, CGDD)` with `SWdw` to look at downwelling shortwave radiation.

**Part 2 on next slide**  

---
**Part 2: Download CHIRPs daily rainfall data for Sutton**

Using this [code](https://code.earthengine.google.com)

- Change the dates for download to start on 5/1/2018 and end on 9/1/2018 
- Export it to your Google Drive account
- Study the code so that you understand how it works
    - Be ready to explain how you would change it to download more than one year's worth of data. 










